---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nlme)
```

# Problem 2

```{r}
dental <- read.table("./HW6-dental.txt", header = TRUE) %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = as.factor(gender)
  )
```

## Part A: Make a spaghetti plot to infer how distance is affected by age and gender.

```{r}
ggplot(dental, 
       aes(age, distance, group = child, col = gender)) +
  geom_line()
```

## Part B: On paper

## Part C: Consider the following marginal model: E(Yij)=β0 +β1 ∗sexi +β2 ∗ageij V ar(Yi) = Σ. Fit a model with (a) compound symmetry covariance, (b) exponential co- variance, (c) autoregressive covariance. Compare the coefficient parameter estimates and the covariance estimates.

### We assume that variances of the response are not constant across age.

```{r}
# Compound Symmetry Covariance

comsym <- gls(distance ~ gender + age, dental, correlation = corCompSymm(form = ~ 1 | child), 
              weights = varIdent(form = ~ 1 | age), method = "REML")

summary(comsym)

# Correlation:
temp_comsym <- corMatrix(comsym$modelStruct$corStruct)[[1]]

# Ratio Between the Variances
ratio_comsym <- 1/varWeights(comsym$modelStruct)[1:4]

# Covariance:
cov_1_comsym <- diag(ratio) %*% temp %*% diag(ratio) * comsym$sigma^2
```

```{r}
# Exponential Covariance
expo <- gls(distance ~ gender + age, dental, correlation = corExp(form = ~ 1 | child),
            weights = varIdent(form = ~ 1 | age), method = "REML")

summary(expo)

#correlation:
temp_expo <- corMatrix(expo$modelStruct$corStruct)[[1]]

# ratio between the variances
ratio_expo <- 1/varWeights(expo$modelStruct)[1:4]

# covariance:
cov_2_expo <- diag(ratio_expo) %*% temp_expo %*% diag(ratio_expo) * expo$sigma^2
```

```{r}
# Autoregressive Covariance
auto1 <- gls(distance ~ gender + age, dental, correlation = corAR1(form = ~ 1 | child), weights = varIdent(form = ~1 | age), method = "REML")

summary(auto1)

#correlation:
temp_auto <- corMatrix(auto1$modelStruct$corStruct)[[1]]

# ratio between the variances
ratio_auto <- 1/varWeights(auto1$modelStruct)[1:4]

# covariance:
cov_3_auto <- diag(ratio_auto) %*% temp_auto %*% diag(ratio_auto) * auto1$sigma^2
result1 <- rbind(round(comsym$coefficients,2),
round(expo$coefficients,2),
round(auto1$coefficients,2))
rownames(result1) <- c("Compound symmetry", "Exponential", "Autoregressive")
```

### Assume that variance of response is constant across age

```{r}
# Compound Symmetry Covariance

comsym <- gls(distance ~ gender + age, dental, correlation = corCompSymm(form = ~ 1 | child), method = "REML")

summary(comsym)

# covariance:
cov_1a <- comsym$sigma^2*corMatrix(comsym$modelStruct$corStruct)[[1]]
```

```{r}
# Exponential Covariance
expo <- gls(distance ~ gender + age, dental, correlation = corExp(form = ~1 | child), method = "REML")
            
summary(expo)
corMatrix(expo$modelStruct$corStruct)[[1]]
# covariance:
cov_2a <- expo$sigma^2*corMatrix(expo$modelStruct$corStruct)[[1]]
```

```{r}
# Autoregressive Covariance

auto1 <- gls(distance ~ gender + age, dental, correlation = corAR1(form = ~ 1 | child), method = "REML")

summary(auto1)
corMatrix(auto1$modelStruct$corStruct)[[1]]

# covariance:
cov_3a <- auto1$sigma^2*corMatrix(auto1$modelStruct$corStruct)[[1]]
result2 <- rbind(round(comsym$coefficients,2),
round(expo$coefficients,2),
round(auto1$coefficients,2))
rownames(result1) <- c("Compound symmetry", "Exponential", "Autoregressive")
```

